Stack Safety as a Security Property

### 2020-07-06 main.tex
Finished examples for lazy properties, suggested slightly different coq for prefix

\begin{figure}
\[
\begin{array}{c|c|c|c|c}
  \row{\sw ~ \rsp ~ r_4 ~ \negate 2}{[\rsp - 2] \leftarrow r_4}{[107] \leftarrow 18}
      {Store $r_4$ as the return value}
  \row{\lw ~ \rra ~ \rsp ~ \negate 1}{\rra \leftarrow [\rsp - 1]}{\rra \leftarrow 31}
      {Load return address}
  \row{\addi ~ \rsp ~ \rsp ~ \negate 1}{\rsp \leftarrow \rsp - 1}{\rsp \leftarrow 108}
      {Restore the stack pointer of {\tt g}}
  \row{\jalr ~ \rra ~ \rra ~ 0}{\leo{??}}{\PCname \leftarrow 31}
      {Return to {\tt g}}
  \end{array}
\]

@@

@@

A trace over a type $A$ is a nonempty, finite or infinite sequence of
$A$s. We will mostly be concerned with sequences of pairs of machine states
and observations, called machine traces and denoted $\machOT \in \MOTRACE$,
and with sequences of triples of a machine state, a policy state, and an
observation, denoted $\MPT \in \MPOTRACE$ and referred to as MP-traces. 
\rb{The way the definitions are set up, one may expect MPO, but\ldots}

@@

The ``trace-of'' operator, written \(\hookrightarrow\), coinductively
relates a machine state with the trace of machine states paired with
observations produced by repeated application of step:
%
\[\mach \hookrightarrow \machOT : \MACHS \rightarrow \MOTRACE\]%
%
\judgmenttwo{\(\mach_0 \stepstoobs{\obs} \mach_1\)}{\(\mach_1 \hookrightarrow \machOT\)}
            {\(\mach_0 \hookrightarrow (\mach_0,\obs) \machOT\)}%
%
\noindent
Since the step function is total, if \(\mach \hookrightarrow \machOT\),
then \(\machOT\) must be infinite.

@@

In a similar fashion, we define \(\mpstatename \hookrightarrow_P \MPT\)
to relate a machine-policy state pair to the MP-trace
induced by by \(\stepstoobs{}_P\). Unlike the $\MOTRACE$ above, this trace can
be finite as the policy can cause a fault, in which case the final state is
a \(\tau\):
%
\judgment{\(\neg \exists \pol_1 . \pol_0 \rightharpoonup \pol_1\)}
         {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow \mpostate{\mach_0}{\pol_0}{\tau}\)}
%
\judgmentthree{\(\mach_0 \stepstoobs{\obs} \mach_1\)}
              {\(\pol_0 \rightharpoonup \pol_1\)}
              {\(\mpstate{\mach_1}{\pol_1} \hookrightarrow \machOT\)}
              {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \machOT\)}
We will also use \(\pi_m\) to project the trace of machine states
out of a trace, \(\pi_p\) to project the trace of policy states, and
\(\pi_o\) to project the observations, with \(\pi_m\) and \(\pi_o\) overloaded
for machine traces and MP-traces.

@@

\judgmenttwo{\(\mach \hookrightarrow \machOT'\)}
            {\(\PUT(\lambda (\mach',\obs).f\ \mach',\machOT')
              = \machOT\)}
            {\(\mach \hookrightarrow \machOT | f\)}
\noindent
And similarly for traces with policies:

\judgmenttwo{\(\mpstatename \hookrightarrow \MPT'\)}
            {\(\PUT(\lambda \mpostate{\mach}{\pol}{\obs}.f(\mach),\MPT') = \MPT\)}
         {\(\mpstatename \hookrightarrow \MPT | f\)}

@@

@@

\definition{Transition}
Given a trace $MP$, a call map $\callmap$ and a code map $\codemap$,
an $n$-{\em transition} to a callee $f$ is a pair of machine-policy
state pairs $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$ in the trace
such that:
\begin{itemize}
\item one state pair steps to the other:
  $$\exists \obs. ~ (\mach_c, \pol_c) \stepstoobs{\obs}_P (\mach_e, \pol_e),$$
\item $\mach_c$ is a valid call with $n$ arguments:
  $$\callmap(\mach_c(\PCname)) = n,$$
\item and $\mach_e$ is an entry point for $f$:
  $$\codemap(\mach_e(\PCname)) = f.$$
\end{itemize}

For an example see \cref{fig:intex}, in which {\tt h} writes its return
value to {\tt g}'s return location the stack rather than its own. That address
is high integrity (red), and so writing to it causes the property to fail.

@@

\Cref{fig:confex2} shows the initial states \(\mach_e\), left,
and \(\mach_e'\), right. Low confidentiality data in green is shared
between them, while high confidentiality data in red and yellow may differ.
After executing the code shown, the final states below have written different
values to address 42, represented by the black and white components.
It is allowed for address 40 to differ between the states due to the initial
variation, but address 42 changed differently between the traces, representing
a leak of data. %Meanwhile \ref{fig:lazyex2}, while high confidentiality data
%is read, it is overwritten before return, and the write at step 3 always writes
%the same value. Yellow and red locations differ between variations only because
%of the initial variation, and the rest are identical. So this example passes.

\begin{figure}
  \confidentialityduringexample
  \caption{Failing Trace Confidentiality by Varying Observations}
  \label{fig:confex1}
\end{figure}

\begin{figure}
  \confidentialityendexample
  \caption{Failing Trace Confidentiality at End State}
  \label{fig:confex2}
\end{figure}

\Cref{fig:confex2} gives an example of a trace that fails trace
confidentiality. At instruction 42, {\tt h} reads {\tt x} from {\tt main}.
Since {\tt x}, at address 100, is high confidentiality (red) the variant
may have a different value (yellow). So both \(r_4\) and address 107
differ in the final state between the original trace and the variant one,
failing the ending condition of trace confidentiality.

@@

@@

      In \cref{fig:lazyex1}, the final states of the original and variant
      differ, because a written value (in black or white) depended on the value
      of a high confidentiality location. So the resulting trace fails trace
      confidentiality. For observable trace confidentiality, we ask whether the
      differing values at address 42 cause any observable distinctions in
      future behavior. To do so, we must undo the variation of formerly high
      confidentiality locations, so that those in yellow match those in red,
      while preserving the updated values where data may have been leaked.

@@

>>>>

\begin{figure}
\[
\begin{array}{c|c|c|c|c}
  \row{\sw ~ \rsp ~ r_4 ~ \negate 2}{[\rsp - 2] \leftarrow r_4}{[107] \leftarrow 18}
      {Store $r_4$ as the return value}
  \row{\lw ~ \rra ~ \rsp ~ \negate 1}{\rra \leftarrow [\rsp - 1]}{\rra \leftarrow 27}
      {Load return address}
  \row{\addi ~ \rsp ~ \rsp ~ \negate 1}{\rsp \leftarrow \rsp - 1}{\rsp \leftarrow 108}
      {Restore the stack pointer of {\tt g}}
  \row{\jalr ~ \rra ~ \rra ~ 0}{\leo{??}}{\PCname \leftarrow 27}
      {Return to {\tt g}}
  \end{array}
\]

@@

\sna{Variations here collides with state variations. I propose substitutions.}

@@

A trace over a type $\TRACEELEMS$ is a nonempty, finite or infinite sequence of
$\TRACEELEMS$s. We will mostly be concerned with sequences of pairs of machine states
and observations, called machine traces and denoted $\machOT \in \MOTRACE$,
and with sequences of triples of a machine state, a policy state, and an
observation, denoted $\MPT \in \MPOTRACE$ and referred to as MP-traces. 
\rb{The way the definitions are set up, one may expect MPO, but\ldots}

@@

The ``trace-of'' operator, written \(\mach \hookrightarrow \machOT\),
coinductively relates a machine state with the trace of machine states
paired with observations produced by repeated application of step:

\judgmenttwo{\(\mach_0 \stepstoobs{\obs} \mach_1\)}{\(\mach_1 \hookrightarrow \machOT\)}
            {\(\mach_0 \hookrightarrow (\mach_0,\obs) \machOT\)}%
%
\noindent
Since the step function is total, if \(\mach \hookrightarrow \machOT\),
then \(\machOT\) must be infinite.

@@

In a similar fashion, we define \(\mpstatename \hookrightarrow_P \MPT\)
to relate a machine-policy state pair to the MP-trace
induced by by \(\stepstoobs{}_P\). Unlike the $\MOTRACE$ above, this trace can
be finite as the policy can cause a fault, in which case the final state is
a \(\tau\):

\begin{minipage}{.4\textwidth}
\judgment{\(\neg \exists \pol_1 . \pol_0 \rightharpoonup \pol_1\)}
         {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow \mpostate{\mach_0}{\pol_0}{\tau}\)}
\end{minipage}
\begin{minipage}{.55\textwidth}
\judgmentthree{\(\mach_0 \stepstoobs{\obs} \mach_1\)}
              {\(\pol_0 \rightharpoonup \pol_1\)}
              {\(\mpstate{\mach_1}{\pol_1} \hookrightarrow \machOT\)}
              {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \machOT\)}
\end{minipage}

We will also use \(\pi_m\) to project the trace of machine states
out of a trace, \(\pi_p\) to project the trace of policy states, and
\(\pi_o\) to project the observations, with \(\pi_m\) and \(\pi_o\) overloaded
for machine traces and MP-traces.

@@

\judgmenttwo{\(\mach \hookrightarrow \machOT'\)}
            {\(\PUT(f,\machOT')
              = \machOT\)}
            {\(\mach \hookrightarrow \machOT | f\)}
\noindent
And similarly for traces with policies:

\judgmenttwo{\(\mpstatename \hookrightarrow \MPT'\)}
            {\(\PUT(\lambda \mpstate{\mach}{\pol}.f\ \mach,\MPT') = \MPT\)}
         {\(\mpstatename \hookrightarrow \MPT | f\)}

@@

For an example see \cref{fig:intex}, representing the unsafe substitution
\[
  44: ~ \sw ~ \rsp ~ r_4 ~ -5
\]
This instruction is in {\tt h}, and instead of storing data to the return,
it overwrites {\tt g}'s variable, {\tt z}. At the top of the figure high
integrity memory is in red, while low integrity is in green; at the end of the
trace address 103 should match the initial state, and does not. So trace
integrity is violated.

@@

\definition{Transition}
Given a trace $\MPT$, a call map $\callmap$ and a code map $\codemap$,
an $n$-{\em transition} to a callee $f$ is a pair of machine-policy
state pairs $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$ in the trace
such that:
\begin{itemize}
\item one state pair steps to the other:
  $$\exists \obs. ~ (\mach_c, \pol_c) \stepstoobs{\obs}_P (\mach_e, \pol_e),$$
\item $\mach_c$ is a valid call with $n$ arguments:
  $$\callmap(\mach_c(\PCname)) = n,$$
\item and $\mach_e$ is an entry point for $f$:
  $$\codemap(\mach_e(\PCname)) = f.$$
\end{itemize}

@@

For example, consider the unsafe substitution:
\[
42 : ~ \sw ~ r_4 ~ \rsp 100
\]
Instead of reading the proper argument, {\tt h} reads directly from the
secret value of {\tt x} that belongs to {\tt main}. \Cref{fig:confex}
shows \(\mach_e\), the entry state of {\tt h}, with high confidentiality memory
in red and low confidentiality in green. Below it, the variant state
\(\mach_e'\) agrees on low confidentiality memory, but the rest might have
any value, in yellow. After the last step, the return states \(\mach_r\) and
\(\mach_r'\) should agree on every component that changed since \(\mach_e\) or
\(\mach_e'\). Addresses 108 and 109 have changed, shown in pink, and while 109
is consistent, 108 will not hold 43 in all variants, so the trace does not obey
trace confidentiality.

Meanwhile, the following unsafe substitution outputs the secret
directly, violating the internal observation clause of confidentiality:
\[
30 : ~ \lw ~ r_5 ~ \rsp ~ \negate 6
\]
%\Cref{fig:confex2} shows the initial states \(\mach_e\), above,
%and \(\mach_e'\), below. Low confidentiality data in green is shared
%between them, while high confidentiality data in red and yellow may differ.
%After executing the code shown, the final states below have written different
%values to address 42, represented by the black and white components.
%It is allowed for address 40 to differ between the states due to the initial
%variation, but address 42 changed differently between the traces, representing
%a leak of data. %Meanwhile \ref{fig:lazyex2}, while high confidentiality data
%is read, it is overwritten before return, and the write at step 3 always writes
%the same value. Yellow and red locations differ between variations only because
%of the initial variation, and the rest are identical. So this example passes.

%\begin{figure}
%  \confidentialityduringexample
%  \caption{Failing Trace Confidentiality by Varying Observations}
%  \label{fig:confex1}
%\end{figure}

\begin{figure}
  \confidentialityendexample
  \caption{Failing Trace Confidentiality}
  \label{fig:confex}
\end{figure}

%\Cref{fig:confex2} gives an example of a trace that fails trace
%confidentiality. At instruction 42, {\tt h} reads {\tt x} from {\tt main}.
%Since {\tt x}, at address 100, is high confidentiality (red) the variant
%may have a different value (yellow). So both \(r_4\) and address 107
%differ in the final state between the original trace and the variant one,
%failing the ending condition of trace confidentiality.

@@

    For out example, we return to the violation of trace integrity in
    \cref{fig:intex}, in which {\tt h} has written to {\tt z} and returned.
    Whether observable integrity holds depends on the continuing trace, shown
    in \cref{fig:intlex}. Above, we have the final state of \(\MPT\),
    \(\mach_r\), with high integrity components in red and low integrity in
    green. Next is the rolled-back state, \(\mach_r' =
    \mathit{roll}(\contour,\mach_e,\mach_r)\), in which high integrity
    components are restored to their value as of \(\mach_e\). Specifically,
    address 103 is restored to 0. When instruction 31 prints {\tt z} in the
    trace from \(\mach_r\), it prints 18, but in the trace from \(\mach_r'\)
    it prints 0. So observable trace integrity does not hold.

    However, if {\tt g} had an additional instruction to overwrite {\tt z}
    before printing, observable trace integrity would hold.

@@

      To return to the example in \cref{fig:confex}, the varied final state
      \(\mach_r'\) might do anything depending on the values of the variation,
      so we cannot immediately compare traces from \(\mach_r\) and
      \(\mach_r'\). We must first restore the varied values to their original
      state, while preserving differences that have occurred during
      execution.

%      In \cref{fig:lazyex1}, the final states of the original and variant
%      differ, because a written value (in black or white) depended on the value
%      of a high confidentiality location. So the resulting trace fails trace
%      confidentiality. For observable trace confidentiality, we ask whether the
%      differing values at address 42 cause any observable distinctions in
%      future behavior. To do so, we must undo the variation of formerly high
%      confidentiality locations, so that those in yellow match those in red,
%      while preserving the updated values where data may have been leaked.

@@

      \begin{figure}
        \confidentialitylazyexample
        \caption{Continuing from \cref{fig:confex}}
        \label{fig:conflex}
      \end{figure}

      In our example, \cref{fig:conflex} shows how execution continues from
      \(\mach_r'\) and the restored \(\mach_r'' = \mathit{restore}(\contour,
      \mach_e,\mach_e',\mach_r,\mach_r'\). Addresses 108 and 109, in pink, are
      not restored; every address that was varied and not changed in one of the
      executions is. Since 108 contains a value that originates in the
      variation, values that depend on it might be anything; after returning
      from {\tt g}, {\tt main} prints 95 in the trace from \(\mach_r\) and
      some undetermined value in the trace from \(\mach_r''\). So observable
      trace confidentiality does not hold.
